\huge{ \emph{Supplement to} Algorithms For Shaping a Particle Swarm\\ With a Shared Control Input Using Boundary Interaction}

### 2016-09-08 RA-L/experiment.tex
Edits made to the paper

\begin{figure}
\begin{center}
	\includegraphics[width=.9\columnwidth]{SetUp.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:setup}
Hardware platform:  table with 1$\times$1 m workspace, surrounded by eight  triggered LED floodlights, with an overhead machine vision system.
}
\vspace{-1em}
\end{figure}
\subsection{Hardware Experiment: Position Control of Two Robots}
The walls of the hardware platform have almost infinite friction, due to a laser-cut, zigzag border. When a kilobot is steered into the zigzag border, they pin themselves to the wall unless the global input directs them away from the wall.  This wall friction is sufficient to enable independent control of two kilobots, as shown in Fig.~\ref{fig:storyReal}.

\subsection{Hardware Experiment: Position Control of n Robots}
The hardware setup has a bounded platform, magnetic sliders, and a magnetic guide board.  Designs for each are available at \citep{arunhardware}. The pink boundary is toothed with a white free space, as shown in Fig \ref{fig:construction2d}. Only discrete, 1 cm moves in the $x$ and $y$ directions are used. The goal configuration highlighted in the top right corner represents a `U' made of seven sliders. The dark red configuration is the current position of the sliders. 
Due to the discretized movements allowed by the boundary, drift moves follow a 1 cm square.  Free robots return to their start positions but robots on the boundary to move laterally, generating a net sliding motion.

@@

\subsection{Hardware Experiment: Control of Covariance}
To demonstrate covariance control $n$=100 robots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.  The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm.

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{newExperiment.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:covExperiment}
Hardware demonstration steering 100 kilobot robots to desired covariance. The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds. The actual covariance is shown in different trials. Frames above the plot show output from machine vision system and an overlaid covariance ellipse.
\vspace{-1em}
}
\end{figure}

>>>>

\begin{figure}
\begin{center}
	\includegraphics[width=.9\columnwidth]{SetUp.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:setup}
Hardware platform:  table with 1$\times$1 m workspace, surrounded by eight  triggered LED floodlights, with an overhead machine vision system.
}
\vspace{-1em}
\end{figure}
\subsection{Hardware Experiment: Position Control of Two Robots}
The walls of the hardware platform have almost infinite friction, due to a laser-cut, slotted border. When a kilobot is steered into the slotted border, they pin themselves to the wall unless the global input directs them away from the wall.  Even though the kilobots are stochastic, this wall friction is sufficient to enable independent control of two kilobots, as shown in Fig.~\ref{fig:storyReal}.

\subsection{Hardware Experiment: Position Control of n Robots}
Kilobots have too much stochasticity to implement Alg. \ref{alg:PosControlNRobots}. Instead, a hardware setup with a bounded platform, magnetic sliders, and a magnetic guide board was used.  Designs for each are available at \citep{arunhardware}. The pink boundary is toothed with a white free space, as shown in Fig \ref{fig:construction2d}. Only discrete, 1 cm moves in the $x$ and $y$ directions are used. The goal configuration highlighted in the top right corner represents a `U' made of seven sliders. The dark red configuration is the current position of the sliders. 
Due to the discretized movements allowed by the boundary, drift moves follow a 1 cm square.  Free robots return to their start positions but robots on the boundary to move laterally, generating a net sliding motion.

@@

\subsection{Hardware Experiment: Control of Covariance}
To demonstrate covariance control of up to 100 kilobots were placed on the workspace and manually steered with lights, using friction with the boundary walls to vary the covariance from  -4000 to 3000 cm$^2$.  The resulting covariance is plotted in Fig.~\ref{fig:covExperiment}, along with snapshots of the swarm.

\begin{figure}
\begin{center}
	\includegraphics[width=\columnwidth]{newExperiment.pdf}
\end{center}
\vspace{-1em}
\caption{\label{fig:covExperiment}
Hardware demonstration steering $\approx 50$ kilobot robots to desired covariance. The goal covariance is negative in first 100 seconds and is positive in the next 100 seconds. The actual covariance is shown in different trials. Frames above the plot show output from machine vision system and an overlaid covariance ellipse.
\vspace{-1em}
}
\end{figure}

### 2016-09-08 RA-L/intro.tex
Edits made to the paper

\section{Introduction}\label{sec:Intro}
Particle swarms steered by a global force are common in applied mathematics, biology, and computer graphics. 
\begin{align}
[\dot{x}_i, \dot{y}_i]^\top = [u_x, u_y]^\top, \qquad i \in [1,n] \label{eq:swarmDynamics}
\end{align}
The control problem is to design $u_x(t), u_y(t)$ to make all $n$ particles achieve a task.
As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015}, \citep{martel2014computer}, \citep{kim2015imparting}, \citep{Donald2013}, \citep{Ghosh2009}, \citep{Ou2013} or \citep{qiu2015magnetic}.
Someday large swarms of robots will be remotely guided
  ex vivo to assemble structures in parallel and 
 through the human body, to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
 For each application, large numbers of micro robots are required to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a global, broadcast signal. 
These applications require techniques to shape the swarm that can reliably exploit large populations despite high under-actuation.

@@

%One surprising result was that humans that only knew the swarm's mean and covariance completed the task faster that humans who knew the position of every robot~\citep{Becker2013b}. Our previous work focused on a block-pushing task, where a swarm of robots pushed a larger block through a 2D maze. 
Even without obstacles or boundaries, the mean position of the swarm in \eqref{eq:swarmDynamics} is controllable.  By adding rectangular boundary walls, some higher-order moments such as the swarm's position variance orthogonal to the boundary walls ($\sigma_x$ and $\sigma_y$ for a workspace with axis-aligned walls) are also controllable~\citep{ShahrokhiIROS2015}. 
A limitation is that global control can only compress a swarm orthogonal to obstacles.  However, navigating through narrow passages often requires control of the variance and the covariance.

The paper is arranged as follows. \S\ref{subsec:FluidInTank} provides analytical position control results in two canonical workspaces with frictionless walls.  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes, \S \ref{subsec:WallFriction} introduces wall friction to the system model.  We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots in \S \ref{sec:PostionControl2Robots}, and \S \ref{sec:PostionControlnRobots} extends this to prove a rectangular workspace with high-friction boundaries can position a swarm of $n$ robots arbitrarily within a subset of the workspace.
\S \ref{sec:simulation} describes implementations of both position control algorithms in simulation and  \S \ref{sec:expResults} describes experiments with a hardware setup and up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}. After a review of recent related work \S \ref{sec:RelatedWork}, we end with directions for future research \S \ref{sec:conclusion}.
%For controlling $\sigma_{xy}$, we prove that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction. 
%We then prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots. 
%We conclude by designing controllers that efficiently regulate $\sigma_{xy}$.
%This paper
%(1) proves that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction,  %where do we do this?
%(2) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots, 
%(3) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots, 
%(4) shows full-state position control with 2 or more robots using  extensive simulations, and
%(5) demonstrate covariance control on our hardware platform with a large number of hardware robots.
%TODO JOURNAL: design controllers that efficiently regulate $\sigma_{xy}$.
%TODO JOURNAL: We will design Lyapunov-inspired controllers for $\sigma_{xy}$ to prove controllability. 
%TODO JOURNAL:  and rank controllability as a function of friction.
% TODO: JOURNAL: and vary wall friction by laser-cutting boundary walls with a variety of profiles. 
%\begin{figure}[t]
%\centering
%\begin{overpic}[width = \columnwidth]{Covariance.jpg}\end{overpic}
%\vspace{-1em}
%\caption{\label{fig:covFriction} Maintaining group cohesion while steering a swarm through an arbitrary maze requires covariance control.
%}\vspace{-1em}
%\end{figure}
% Our paper is organized as follows.  After a discussion of related work in Section \ref{sec:RelatedWork}, we describe our experimental methods for an online human-user experiment in Section \ref{sec:expMethods}.  We report the results of our experiments in Section \ref{sec:expResults}, discuss the lessons learned in Section \ref{sec:discussion}, and end with concluding remarks in Section \ref{sec:conclusion}.

>>>>

\section{Introduction}\label{sec:Intro}
Particle swarms steered by a global force are common in applied mathematics, biology, and computer graphics. 
\begin{align}
[\dot{x}_i, \dot{y}_i]^\top = [u_x, u_y]^\top, \qquad i \in [1,n] \label{eq:swarmDynamics}
\end{align}
The control problem is to design $u_x(t), u_y(t)$ to make all $n$ particles achieve a task.
As a current example, micro- and nano-robots can be manufactured in large numbers, see~\citep{Chowdhury2015}, \citep{martel2014computer}, \citep{kim2015imparting}, \citep{Donald2013}, \citep{Ghosh2009}, \citep{Ou2013} or \citep{qiu2015magnetic}.
Someday large swarms of robots will be remotely guided
  ex vivo to assemble structures in parallel and 
 through the human body to cure disease, heal tissue, and prevent infection. %Both tasks are high   as described in \citep{munoz2014review} \citep{sitti2015biomedical} . 
 For each application, large numbers of micro robots are required to deliver sufficient payloads, but the small size of these robots makes it difficult to perform onboard computation.  Instead, these robots are often controlled by a global, broadcast signal. 
These applications require techniques to shape the swarm that can reliably exploit large populations despite high under-actuation.

@@

%One surprising result was that humans that only knew the swarm's mean and covariance completed the task faster that humans who knew the position of every robot~\citep{Becker2013b}. Our previous work focused on a block-pushing task, where a swarm of robots pushed a larger block through a 2D maze. 
Even without obstacles or boundaries, the mean position of the swarm in \eqref{eq:swarmDynamics} is controllable.  By adding rectangular boundary walls, some higher-order moments, such as the swarm's position variance orthogonal to the boundary walls ($\sigma_x$ and $\sigma_y$ for a workspace with axis-aligned walls), are also controllable~\citep{ShahrokhiIROS2015}. 
A limitation is that global control can only compress a swarm orthogonal to obstacles, but navigating through narrow passages often requires control of the variance and the covariance of the swarm.

The paper is arranged as follows. First \S\ref{subsec:FluidInTank} provides analytical position control results in two canonical workspaces with frictionless walls.  These results are limited in the set of shapes that can be generated.  To extend the range of possible shapes, \S \ref{subsec:WallFriction} introduces wall friction to the system model.  We prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots in \S \ref{sec:PostionControl2Robots}, and \S \ref{sec:PostionControlnRobots} extends this to prove a rectangular workspace with high-friction boundaries can position a swarm of $n$ robots arbitrarily within a subset of the workspace.
\S \ref{sec:simulation} describes implementations of both position control algorithms in simulation and  \S \ref{sec:expResults} describes experiments with a hardware setup and up to 100 robots, as shown in Fig.~\ref{fig:IntroPic}. After a review of recent related work \S \ref{sec:RelatedWork}, we end by suggesting directions for future research in \S \ref{sec:conclusion}.
%For controlling $\sigma_{xy}$, we prove that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction. 
%We then prove that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots. 
%We conclude by designing controllers that efficiently regulate $\sigma_{xy}$.
%This paper
%(1) proves that the swarm position covariance $\sigma_{xy}$ is controllable given boundaries with non-zero friction,  %where do we do this?
%(2) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position two robots, 
%(3) proves that two orthogonal boundaries with high friction are sufficient to arbitrarily position a swarm of $n$ robots, 
%(4) shows full-state position control with 2 or more robots using  extensive simulations, and
%(5) demonstrate covariance control on our hardware platform with a large number of hardware robots.
%TODO JOURNAL: design controllers that efficiently regulate $\sigma_{xy}$.
%TODO JOURNAL: We will design Lyapunov-inspired controllers for $\sigma_{xy}$ to prove controllability. 
%TODO JOURNAL:  and rank controllability as a function of friction.
% TODO: JOURNAL: and vary wall friction by laser-cutting boundary walls with a variety of profiles. 
%\begin{figure}[t]
%\centering
%\begin{overpic}[width = \columnwidth]{Covariance.jpg}\end{overpic}
%\vspace{-1em}
%\caption{\label{fig:covFriction} Maintaining group cohesion while steering a swarm through an arbitrary maze requires covariance control.
%}\vspace{-1em}
%\end{figure}
% Our paper is organized as follows.  After a discussion of related work in Section \ref{sec:RelatedWork}, we describe our experimental methods for an online human-user experiment in Section \ref{sec:expMethods}.  We report the results of our experiments in Section \ref{sec:expResults}, discuss the lessons learned in Section \ref{sec:discussion}, and end with concluding remarks in Section \ref{sec:conclusion}.
